---
title: "ML Project 1"
output: html_document
date: "2024-02-04"
---

This machine learning project seeks to create a valuable model capable of predicting whether a client will subscribe to a business product offered by a Portuguese bank following a marketing campaign.

Loading the libraries

```{r}
library(tidyverse)
library(tidymodels)
library(gtsummary)
```

The dataset has already been separated into training and test datasets. The dataset was published Prakhar Rathi by on Kaggle.

```{r}
train <- read_csv2("train.csv")
str (train)
head (train)

```

Step 1: Exploratory Data Analysis 

1) Understanding the demographics of the clients


```{r}

# Defining the demographic function f 
# has two input variables, the tibble train data and x ( categorical variable in the train tibble)

f <- function (train, x) {
 dem <- train %>% group_by_at (vars({{x}})) %>% 
  summarise ( total_clients = n(),
              percentage = round ( n()* 100 / nrow(train), 2)) %>%
  arrange (desc(total_clients))
 return(dem)
}
```

```{r}
# Education background 
f (train, education)
```

```{r}
# Marital status proportion
f (train, marital)
```

```{r}
# Job description proportion
f (train, job)
```

How many clients have personal loans?

```{r}
f (train, loan)
```
How many clients have credit in default?

```{r}
f(train, default)
```
How many clients have a housing loan

```{r}
f(train, housing)
```
Campaign contact summary

```{r}
# Total contacts performed during the campaign period

sum (train$campaign)
```
```{r}
# Types of communication methods summary
f(train, contact)
```

```{r}
# Average contact (last contact) duration in minutes per contact method

train %>% group_by (contact) %>% 
  summarise ( total_contacts = sum(campaign),
              ave_contact_duration = round ( mean(duration/60),2)) %>%
  arrange (desc(ave_contact_duration))
```

```{r}
# last contact day of the month
f(train, day)
```
```{r}
# last contact month of the year
f(train, month)
```
```{r}
# Days elapsed since the client's last contact from the last campaign ( -1 indicating that the client was not contacted previously)
f(train, pdays)
```
``` {r}
# number of contacts performed before this campaign 
f(train, previous)
```
```{r}
# outcome of the previous marketing campaign
f(train, poutcome)
```

Understanding the demographic effect on the campaign outcome:

```{r}
# First we get a summary of the campaign outcome
train <- train %>% rename( outcome = y)
f(train, outcome)
```

```{r}
# Understanding the education level of clients that subscribed to the the financial product

train_yes <- train %>% filter (outcome == "yes") ## filtering clients that subscribed

# we also want add the total amount of contacts made and average duration to understand the level of resources used

f2 <- function (train_yes, x1, x2,x3) {
 dem <- train_yes %>% group_by_at (vars({{x1}})) %>% 
  summarise ( total_clients = n(),
              percentage = round ( n()* 100 / nrow(train_yes), 2),
              total_contacts = sum({{x2}}),
              contacts_to_clients_ratio = round (total_contacts / total_clients,1),
              ave_duration_min = round (mean(({{x3}})/60),2)) %>%
  arrange (desc(total_clients))
 return(dem)
}

f2(train_yes, education, campaign, duration)
```

```{r}
# Understanding the marital status of clients that subscribed to the financial product

f2 (train_yes, marital, campaign, duration )
```
```{r}
# Understanding the job description of clients that subscribed to the financial product

f2 (train_yes, job, campaign, duration )
```
```{r}
# Understanding the loan status of clients that subscribed to the financial product

f2 (train_yes, loan, campaign, duration )
```
```{r}
# Understanding the default status of clients that subscribed to the financial product

f2 (train_yes, default, campaign, duration )
```
```{r} 
# Understanding the effect of communication method on the number of clients that subscribed to the financial product

f2 (train_yes, contact, campaign, duration )
```
```{r}
# Understanding the housing status of clients that subscribed to the financial product

f2 (train_yes, housing, campaign, duration )
```
```{r}
# Understanding the effect of outcome of previous campaign on clients that subscribed to the financial product

f2 (train_yes, poutcome, campaign, duration )
```
```{r}
# Evaluating if the yearly average balance affect the campaign outcome

f3 <- function (train, x1,x2) {
 dem <- train %>% group_by_at (vars({{x1}})) %>% 
  summarise ( total_clients = n(),
              percentage = round ( n()* 100 / nrow(train), 2),
              ave_balance = round (mean({{x2}}),1)) %>%
  arrange (desc(total_clients))
 return(dem)
}
f3(train, outcome, balance)
```

````{r}
# Education bargraph

train %>% ggplot ( aes(education)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= -0.5, size = 3) +
  theme_classic() + labs (x = "Education level", y = "No. of clients")
```
```{r}
# Marital status bargraph
# Education bargraph

train %>% ggplot ( aes(marital)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= -0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", X = "Marital status")
```
```{r}
# Job description bargraph
train %>% ggplot ( aes(y = job)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",hjust= -0.5, size = 2) +
  theme_classic() + labs (y = "Job description", x = "No. of clients")
```
```{r}
# Job description bargraph
train %>% ggplot ( aes(contact)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= - 0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", x = "Communication method")
```
```{r}
# Housing loan status bargraph 
train %>% ggplot ( aes(housing)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= - 0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", x = "Communication method")
```

```{r}
# Loan status bargraph 
train %>% ggplot ( aes(loan)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= - 0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", x = "Loan status")
```

```{r}
# Loan default status bargraph 
train %>% ggplot ( aes(default)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= - 0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", x = "Loan default status")
```

```{r}
# Previous campaign outcome  bargraph 
train %>% ggplot ( aes(poutcome)) + geom_bar(aes(fill = outcome)) +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "stack",vjust= - 0.5, size = 3) +
  theme_classic() + labs (y = "No. of clients", x = "Previous outcome")
```













